Data Flow Analysis\\in the Presence of Correlated Calls

### 2014-12-05 content/background.tex
restructuring

\subsection{Related Work}
\otodo{should we move Related Work to near the end of the paper?}
IFDS is a version of the functional approach to data-flow analysis developed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}. Their algorithm is based on computing \textit{summary functions} that return the data-flow value at the end of a procedure, given the data-flow value at the start of the procedure. IFDS problems form a more restricted set of data-flow problems: unlike in the functional approach, IFDS flow functions have to be distributive, and the set of data-flow facts $D$ has to be finite. However, the IFDS algorithm is more general than Sharir's and Pnueli's algorithm in that it can handle programs containing local variables and parameters in recursive methods.

IFDS has been used to encode a variety of data-flow problems. More complex examples of applications include typestate analysis (determining which operations can be performed on an object at a given program point)~\cite{naeem2008typestate} or shape analysis (detecting errors and validating properties of programs at compile time)~\cite{DBLP:conf/birthday/KreikerRRSWY13}.

IFDS is implemented for two popular static-analysis frameworks, the T.J. Watson Libraries for Analysis (WALA)~\cite{fink2012wala} and Soot~\cite{vallee1999soot}.

WALA is a framework for static analysis on Java bytecode developed by the IBM T.J.~Watson Research Center.
In the implementation of our work, we use WALA to build and traverse the supergraph (a special kind of control-flow graph) of a Java program\footnote{However, we do not use WALA's IFDS implementation, as explained in~Section~\ref{chapter:eval}.}.

Soot is a framework for program analysis and optimization on Java bytecode, developed by the Sable Research Group at McGill University. Unlike WALA, Soot also has an implementation of the IDE algorithm. The IFDS and IDE implementations for Soot are part of the Heros project~\cite{bodden2012inter}.

Whereas one advantage of Soot's IFDS implementation (and other static analysis tools) is ease of use and extensibility, WALA's primary focus is efficiency. For example, WALA uses bit-vectors to represent some of the analysis data types, like local variables and parameters. Another difference is that WALA's intermediate representation of a program uses static single assignment (SSA) form~\cite{cytron1991efficiently}. SSA form is a representation of the program in which each variable has only one definition (assignment). SSA can make dataflow analysis simpler and more efficient~\cite{appel1998ssa}.

Work on improving the IFDS algorithm includes Practical Extensions by~N.\,Naeem and O.\,Lhot\'ak~\cite{naeem2010pei}. Their paper presents four extensions to the IFDS algorithm. Two of the extensions improve the efficiency of the IFDS analysis for certain classes of IFDS problems. Another extension widens the class of problems applicable for the IFDS analysis. However, those extensions do not affect the precision of IFDS problems. Our analysis, in contrast, does not improve the efficiency or generality of IFDS, but it allows us to solve IFDS problems more precisely.

The fourth extension is targeted towards programs that are represented in SSA form. Executing the IFDS analysis on such programs results in loss of precision in the presence of control-flow constructs (e.g. conditionals and loops), compared to programs in non-SSA form.
The extension makes the IFDS analysis on programs in SSA form as precise as on programs that are not represented in SSA form. In contrast, the correlated-calls analysis is applicable to programs in both SSA and non-SSA forms. Even if applied to a program in SSA form, our analysis and the extension improve the precision of IFDS in unrelated situations: the first analysis handles correlated calls, and the latter handles control-flow constructs. Thus, an IFDS analysis could benefit from both precision improvements independently.

Another work on improving the efficiency of the IFDS algorithm is E.\,Bodden et~al.'s framework for the analysis of software products lines~\cite{bodden2013spl}. Their paper uses transformations from IFDS to IDE problems, a technique we also employ. Finally, J.\,Rodriguez and O.\,Lhot\'ak implemented a concurrent version of the IFDS algorithm using actors~\cite{rodriguez2010concurrent}. However, neither of those works is concerned with improving the precision of IFDS results.

The correlated-calls analysis improves the precision of a data-flow analysis by eliminating a special type of infeasible paths. This is similar to the idea of context-sensitive analysis: just as a context-sensitive analysis eliminates infeasible paths from the end of a procedure to the call sites that do not match the given procedure call, the correlated-calls analysis eliminates infeasible paths caused by correlated method calls.

The idea of using correlated calls to remove infeasible paths in data-flow analyses of object-oriented programs was introduced by F.\,Tip~\cite{DBLP:journals/scp/Tip15}. The possibility of using IDE to achieve this is mentioned, but not elaborated upon. Our work presents a concrete solution to the problem and an implementation of that solution.

The idea of eliminating infeasible paths caused by correlated calls is similar to M.~Sridharan et~al.'s work on improving the precision of pointer analysis for JavaScript programs~\cite{DBLP:conf/ecoop/SridharanDCST12}. For each pointer, a pointer analysis determines the possible set of objects (the \textit{points-to} set) that the pointer can reference at a given program point. In JavaScript, it is challenging to compute the points-to set of fields because in general, field names can be derived from arbitrary expressions and bound at runtime.
As a result, an imprecise data-flow analysis will include infeasible paths between values of the form \verb'o[p]' (access of a property \verb'p' of object \verb'o'), where at compile time, \verb'p' can be bound to different values.
The idea of the paper is to track all dynamic property accesses (reads and writes) on an object \verb'o' with property name \verb'p'. The code snippets containing the references \verb'o[p]' are then extracted into a separate function $f$. The analysis is then run so that for each possible value of \verb'p', $f$ is analyzed separately; therefore, for a given property name, all correlated objects with that name are analyzed together.

The differences between this method of tracking correlated calls and our analysis are the following.
\begin{itemize}
  \item \textit{Type of target data-flow analysis} whose precision is to be improved. Our analysis improves the precision of IFDS data-flow analyses, whereas the JavaScript analysis improves the precision of pointer analysis.
  \item \textit{Target language}. Our analysis is for object-oriented languages where polymorphic methods, and not property names (which are known at compile time), cause infeasible paths.
  \item \textit{Different handling of correlated calls}. Extracting code that contains correlated calls into separate methods would not prevent infeasible paths. Instead, our analysis uses IDE flow functions to detect and eliminate infeasible paths caused by correlated calls.
\end{itemize}

@@

\begin{example}
  Consider the program in Listing~\ref{list:examplejava}. The supergraph corresponding to that program is shown in Figure~\ref{fig:examplesupergraph}.
  \otodo{Can we merge the examples from Figures 1 and 2 into a single example program?}
\otodo{I think we can remove the supergraph Figure 3.}
\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{examplesupergraph.java}
  \end{minipage}
  \caption{An example Java program}
  \label{list:examplejava}
\end{figure}

@@

%arrow description
    \node [description] (intra_arrow_left) [above right = \dist and 2*\dist of st_secret] {};
    \node [description] (intra_arrow_right) [right = \dist of intra_arrow_left] {};
    \node [description] (inter_arrow_left) [below = .5*\dist of intra_arrow_left] {};
    \node [description] (inter_arrow_right) [below = .5*\dist of intra_arrow_right] {};
    \path[->](intra_arrow_left) edge node[right = .5*\dist]{intra-procedural edge} (intra_arrow_right);
    \path[dashed,->](inter_arrow_left) edge node[right = .5*\dist]{inter-procedural edge} (inter_arrow_right);
\end{tikzpicture}
}
  \caption{An example supergraph for Listing~\ref{list:examplejava}}%
  \label{fig:examplesupergraph}%
  \end{figure}
\end{example}

@@

\begin{example}
  In the supergraph in Figure~\ref{fig:examplesupergraph}, let us assign $\{,\,\}$ parentheses to $\textsf{call}_\texttt{A.f}$ and $\textsf{return}_\texttt{A.f}$, and $\langle,\,\rangle$ parentheses to $\textsf{call}_\texttt{f}$ and $\textsf{return}_\texttt{f}$. 
  Then the string corresponding to the path 
  \[
    p_1=\left[\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  is $\{\}$, which indicates that $p_1$ is valid. Every prefix of $p_1$ is also a valid path.

However, the graph also contains an inter-procedurally invalid path
  \[
    p_2=\left[\highlight{\textsf{call}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  with corresponding string $\langle\}$.
\end{example}

@@

@@

In this way, the IFDS algorithm reduces the input data-flow problem to a graph-reachability problem.
\begin{example}
  In a taint analysis, $D$ is the set of variables in the program. If a fact $d\in D$ is reachable at a given node, then the variable is considered secret at that node. Otherwise, it is considered not secret. The question ``is $d$ secret at node $n$?'' becomes ``is there a realizable path from (\textsf{start}$_\texttt{main}$,\,\textbf{0}) to $(n,\,d)$?''.
\end{example}
\begin{example} 
  In an available-expressions analysis, $D$ is the set of all expressions in the program. If an expression $d\in D$ is reachable at a certain node, it means that it does not need to be recomputed at that node. 
\end{example}
\begin{example}
  The exploded supergraph for Listing~\ref{list:examplejava} is shown in Figure~\ref{fig:exampleexploded}.
  We can see that there is a realizable path
  from the start node of the exploded graph to the variable \verb'b' at~the~node~$\textsf{return}_\texttt{A.f}$ in the \verb'main' method. This means that at that node, \verb'b' is considered secret.

    \begin{figure}%
      \hspace*{-30pt}
\scalebox{.9}{
      \tikzset{
        ashadow/.style={
          opacity=.25,
          shadow xshift=0.07,
          shadow yshift=-0.07
        },
      }
    \tikzstyle{supergraph}=[
      fill=greyblue,
      rounded corners,
      font=\small,
      align=right,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_s}=[
      fill=bisque,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_f}=[
      fill=lightsalmonpink,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{description}=[fill=white]
    \tikzstyle{n}=[fill=black,circle,inner sep=1.5pt]

@@

     \node [n,label=$\mathbf0$] (return_x_z) [right=\ndd of return_x.east] {};
     \node [n,label=\texttt{r}] (return_x_r) [right=\ndd of return_x_z] {};
     \node [n,label=\texttt{s}] (return_x_s) [right=\ndd of return_x_r] {};
%f method edges
    \path[->](st_f_z) edge (if_z);
    \path[->](if_z) edge (call_f_f_z);
    \path[->](call_f_f_z) edge (return_r_z);  
    \path[->](return_r_z) edge (end_f_z);
    \path[->](if_z) edge [out=-43,in=127] (return_x_z);
    \path[->](return_x_z) edge [out=270,in=25] (end_f_z);
    \path[->](st_f_s) edge (if_s);
    \path[->](if_s) edge (call_f_f_s);
    \path[->](call_f_f_s) edge (return_r_s);  
    \path[->](return_r_s) edge (end_f_s);
    \path[->](if_s) edge [out=-30,in=110] (return_x_s);
    \path[->](return_x_s) edge [out=270,in=25] (end_f_s);
    \path[->](return_f_f_r) edge (return_r_r);
    \path[->](return_r_r) edge (end_f_r);
    \path[->,dashed](call_f_main_z) edge [out=20,in=160] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=150,in=-20] (return_f_main_z);
    \path[->,dashed](call_f_f_z) edge [out=110,in=250] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=110,in=250] (return_f_f_z);
    \path[->,dashed](call_f_main_x) edge [out=32,in=112] (st_f_s);
    \path[->,dashed](end_f_s) edge [out=155,in=-40] (return_f_main_y);
    \path[->,dashed](end_f_s) edge [out=120,in=-50] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=110,in=250] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=150,in=-50] (return_f_main_y);
\end{tikzpicture}
}
  \caption{The exploded supergraph corresponding to a taint analysis for the input program in Listing~\ref{list:examplejava}}%
  \label{fig:exampleexploded}%
  \end{figure}
\end{example}

@@

\begin{mdelete}
\begin{example}\label{ex:constprop}
  One version of the constant propagation analysis that can be encoded with IDE is \textit{linear constant propagation}. A linear constant propagation analysis can detect constants of the form $a\cdot x+b$, where $a$ and $b$ are integers and $x$ is a variable. In particular, a variable can only be considered constant if it depends on at most one other constant variable: even if $y$ and $z$ are variables that are considered constant, the variable $x=y+z$ will be considered not constant. If we encoded the analysis in a way to handle non-linear constant assignments, we would have to use non-distributive flow functions, which would violate the requirements of the IDE algorithm.

@@

    \path[->] (zero___) edge node[left,sm]{$\id$} (zero____);
    \path[->] (u___) edge node[left,sm]{$\lambda l.5$} (u____);
    \path[->] (v___) edge node[left,sm]{$\id$} (v____);
    \path[->,dashed,color=gray] (w___) edge node[left,sm]{$\id$} (w____);
  \end{tikzpicture}
  \caption[A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}]{A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}. The dashed edges are edges not reachable from the entry node.}%
  \label{fig:constprop}%
  \end{figure}
\end{mdelete}

>>>>

@@

\commentout{
\begin{example}
  Consider the program in Listing~\ref{list:examplejava}. The supergraph corresponding to that program is shown in Figure~\ref{fig:examplesupergraph}.
  \otodo{Can we merge the examples from Figures 1 and 2 into a single example program?}
\otodo{I think we can remove the supergraph Figure 3.}
\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \inputMinted{java}{examplesupergraph.java}
  \end{minipage}
  \caption{An example Java program}
  \label{list:examplejava}
\end{figure}

@@

%arrow description
    \node [description] (intra_arrow_left) [above right = \dist and 2*\dist of st_secret] {};
    \node [description] (intra_arrow_right) [right = \dist of intra_arrow_left] {};
    \node [description] (inter_arrow_left) [below = .5*\dist of intra_arrow_left] {};
    \node [description] (inter_arrow_right) [below = .5*\dist of intra_arrow_right] {};
    \path[->](intra_arrow_left) edge node[right = .5*\dist]{intra-procedural edge} (intra_arrow_right);
    \path[dashed,->](inter_arrow_left) edge node[right = .5*\dist]{inter-procedural edge} (inter_arrow_right);
\end{tikzpicture}
}
  \caption{An example supergraph for Listing~\ref{list:examplejava}}%
  \label{fig:examplesupergraph}%
  \end{figure}
\end{example}
}

@@

\begin{mdelete}
\begin{example}
  In the supergraph in Figure~\ref{fig:examplesupergraph}, let us assign $\{,\,\}$ parentheses to $\textsf{call}_\texttt{A.f}$ and $\textsf{return}_\texttt{A.f}$, and $\langle,\,\rangle$ parentheses to $\textsf{call}_\texttt{f}$ and $\textsf{return}_\texttt{f}$. 
  Then the string corresponding to the path 
  \[
    p_1=\left[\highlight{\textsf{call}_\texttt{A.f}}{greyblue},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  is $\{\}$, which indicates that $p_1$ is valid. Every prefix of $p_1$ is also a valid path.

\mtodo{The figure that we deleted was constructed so that there would be a realizable and a non-realizable path. Now that we deleted that figure, let's delete this example as well.}

However, the graph also contains an inter-procedurally invalid path
  \[
    p_2=\left[\highlight{\textsf{call}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{start}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\texttt{if (s == null)}}{lightsalmonpink},\,
          \highlight{\texttt{return s}}{lightsalmonpink},\,
          \highlight{\textsf{end}_\texttt{f}}{lightsalmonpink},\,
          \highlight{\textsf{return}_\texttt{A.f}}{greyblue}
    \right]
  \]
  with corresponding string $\langle\}$.
\end{example}
\end{mdelete}

@@

\subsection{IFDS and IDE}\mtodo{I will merge this section with the following two sections}
The IFDS framework is a precise and efficient algorithm for data-flow analysis. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}. The IFDS analysis is a version of the classic \textit{functional approach} to data-flow analysis proposed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.

Given a data-flow problem that satisfies the restrictions of IFDS, the algorithm provides a \textit{context-sensitive} solution in polynomial time. In other data-flow algorithms not based on the functional approach, the result of the analysis at the entry of a procedure ``merges'' the incoming data obtained from all callers of the procedure. As a consequence, there is one global data-flow result computed at the end of the procedure. Context-sensitivity, however, allows an analysis to compute the data-flow result for a given procedure as a \textit{function} of the data-flow value at the start of the procedure. 
In other words, the analysis result for a procedure varies depending on where the procedure was called from. This significantly improves the precision of a data-flow analysis, which is why context-sensitivity is an important advantage of IFDS over classic data-flow algorithms.

Compared to IFDS, most data-flow analyses are either general but do not run in polynomial time~\cite{knoop1992interprocedural,pnueli1981two} or handle a very specific set of problems~\cite{knoop1993efficient}.

The IFDS algorithm is applicable to problems which can be expressed with data-flow functions that satisfy certain restrictions. \textit{Inter-procedural} flow functions specify how data flows from the invocation of a procedure to its start, and from the procedure's end back to its call site. \textit{Distributive} flow functions are those that distribute over the confluence operator. In the context of IFDS, the confluence operator is called meet, and it can be either union or intersection. The data-flow facts on which the analysis operates must be a \textit{finite} set $D$. Each flow function operates on a \textit{subset} of $D$ (for example, the set of variables in the program) which makes the domain of the flow functions the power set of $D$. We describe the IFDS restrictions in detail in Section~\ref{sec:suitableifds}.

The IDE framework is an expressive extension to IFDS that was created by the same authors in 1996.
The problems that IDE can solve include, but are not limited to, IFDS problems. Just as the IFDS algorithm, the IDE algorithm is suitable for data-flow analyses that can be encoded with inter-procedural, distributive flow functions. However, in IDE, the domain of the flow functions is not restricted to sets $D$ of data-flow facts. The IDE domain of a flow function consists of \textit{environments} that map data-flow facts from the set $D$ to lattice elements.

As an example, in a constant propagation problem, an IDE environment would map each variable to the (possibly) constant value that it is bound to. To illustrate the distinction between IFDS and IDE we could say that IFDS can find out which variables in a program are constants, whereas IDE can additionally retrieve the values of the constant variables.

@@

In this way, the IFDS algorithm reduces the input data-flow problem to a graph-reachability problem.
\begin{example}
  In a taint analysis, $D$ is the set of variables in the program. If a fact $d\in D$ is reachable at a given node, then the variable is considered secret at that node. Otherwise, it is considered not secret. The question ``is $d$ secret at node $n$?'' becomes ``is there a realizable path from (\textsf{start}$_\texttt{main}$,\,\textbf{0}) to $(n,\,d)$?''.
\end{example}
\begin{example} 
  In an available-expressions analysis, $D$ is the set of all expressions in the program. If an expression $d\in D$ is reachable at a certain node, it means that it does not need to be recomputed at that node. 
\end{example}
\begin{example}
  The exploded supergraph for Listing~\ref{list:ccexample} is shown in Figure~\ref{fig:cc_edgefn_example}.
  We can see that there is a realizable path
  from the start node of the exploded graph to the variable \verb's' at~the~node~$\textsf{print(s)}$ in the \verb'B.bar' method. This means that at that node, \verb's' is considered secret.

\commentout{  
    \begin{figure}%
      \hspace*{-30pt}
\scalebox{.9}{
      \tikzset{
        ashadow/.style={
          opacity=.25,
          shadow xshift=0.07,
          shadow yshift=-0.07
        },
      }
    \tikzstyle{supergraph}=[
      fill=greyblue,
      rounded corners,
      font=\small,
      align=right,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_s}=[
      fill=bisque,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{supergraph_f}=[
      fill=lightsalmonpink,
      rounded corners,
      font=\small,
      align=left,
      text=charcoal,
      drop shadow={ashadow, color=greyblue}
    ]
    \tikzstyle{description}=[fill=white]
    \tikzstyle{n}=[fill=black,circle,inner sep=1.5pt]

@@

     \node [n,label=$\mathbf0$] (return_x_z) [right=\ndd of return_x.east] {};
     \node [n,label=\texttt{r}] (return_x_r) [right=\ndd of return_x_z] {};
     \node [n,label=\texttt{s}] (return_x_s) [right=\ndd of return_x_r] {};
%f method edges
    \path[->](st_f_z) edge (if_z);
    \path[->](if_z) edge (call_f_f_z);
    \path[->](call_f_f_z) edge (return_r_z);  
    \path[->](return_r_z) edge (end_f_z);
    \path[->](if_z) edge [out=-43,in=127] (return_x_z);
    \path[->](return_x_z) edge [out=270,in=25] (end_f_z);
    \path[->](st_f_s) edge (if_s);
    \path[->](if_s) edge (call_f_f_s);
    \path[->](call_f_f_s) edge (return_r_s);  
    \path[->](return_r_s) edge (end_f_s);
    \path[->](if_s) edge [out=-30,in=110] (return_x_s);
    \path[->](return_x_s) edge [out=270,in=25] (end_f_s);
    \path[->](return_f_f_r) edge (return_r_r);
    \path[->](return_r_r) edge (end_f_r);
    \path[->,dashed](call_f_main_z) edge [out=20,in=160] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=150,in=-20] (return_f_main_z);
    \path[->,dashed](call_f_f_z) edge [out=110,in=250] (st_f_z);
    \path[->,dashed](end_f_z) edge [out=110,in=250] (return_f_f_z);
    \path[->,dashed](call_f_main_x) edge [out=32,in=112] (st_f_s);
    \path[->,dashed](end_f_s) edge [out=155,in=-40] (return_f_main_y);
    \path[->,dashed](end_f_s) edge [out=120,in=-50] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=110,in=250] (return_f_f_r);
    \path[->,dashed](end_f_r) edge [out=150,in=-50] (return_f_main_y);
\end{tikzpicture}
}
  \caption{The exploded supergraph corresponding to a taint analysis for the input program in Listing~\ref{list:examplejava}}%
  \label{fig:exampleexploded}%
  \end{figure}
}
\end{example}

@@

\commentout{
\begin{mdelete}
\begin{example}\label{ex:constprop}
  One version of the constant propagation analysis that can be encoded with IDE is \textit{linear constant propagation}. A linear constant propagation analysis can detect constants of the form $a\cdot x+b$, where $a$ and $b$ are integers and $x$ is a variable. In particular, a variable can only be considered constant if it depends on at most one other constant variable: even if $y$ and $z$ are variables that are considered constant, the variable $x=y+z$ will be considered not constant. If we encoded the analysis in a way to handle non-linear constant assignments, we would have to use non-distributive flow functions, which would violate the requirements of the IDE algorithm.

@@

    \path[->] (zero___) edge node[left,sm]{$\id$} (zero____);
    \path[->] (u___) edge node[left,sm]{$\lambda l.5$} (u____);
    \path[->] (v___) edge node[left,sm]{$\id$} (v____);
    \path[->,dashed,color=gray] (w___) edge node[left,sm]{$\id$} (w____);
  \end{tikzpicture}
  \caption[A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}]{A labeled exploded supergraph for a constant-propagation analysis described in Example~\ref{ex:constprop}. The dashed edges are edges not reachable from the entry node.}%
  \label{fig:constprop}%
  \end{figure}
\end{mdelete}
}

### 2014-12-05 content/impl.tex
add comments on what to remove
restructuring

An overview of the transformations is shown in Figure~\ref{fig:transformations}.
\begin{figure}
  \centering
    \tikzset{
  ashadow/.style={opacity=.25, shadow xshift=0.07, shadow yshift=-0.07},
}
  \tikzstyle{problem}=[fill=greyblue,text width=2.3cm,rounded corners,font=\small,text=charcoal,drop shadow={ashadow, color=greyblue}]
  \tikzstyle{result}=[fill=bisque,text width=1.8cm,rounded corners,font=\small,text=charcoal,drop shadow={ashadow, color=greyblue}]
\begin{tikzpicture}
    \node [problem] (ifds) {IFDS problem};
    \node [problem] (equiv) [above right=0.9cm and .7\dist of ifds.east] {Equivalent IDE problem};
    \node [problem] (ccide)[below=3cm of equiv.west, anchor=west] {Correlated-calls IDE problem};
    \node [result] (equivres) [right=\dist of equiv.east] {Equivalence-IDE result};
    \node [result] (ccres) [below=3cm of equivres.west,anchor=west] {Correlated-calls result};
    \node [result] (ifdsres) [right=\dist of equivres.east] {IFDS result};
    \node [result] (improved) [right=\dist of ccres.east] {Improved IFDS result};
    \path[->] (equivres) edge node[above]{$\backEq$} (ifdsres);
    \path[->] (ccres) edge node[above] {$\backCC$} (improved);
    \path[->] (ccres) edge[out=30,in=240] node[above] {$\backEq$} (ifdsres);
    \path[->] (ifds) edge[out=40,in=190] node[above] {$\transEq$} (equiv);
    \path[->] (ifds) edge[out=-60,in=165] node[above] {$\transCC$} (ccide);
    \path[->] (equiv) edge node[above]{$\result$} (equivres);
    \path[->] (ccide) edge node[above]{$\result$} (ccres);
\end{tikzpicture}
  \caption{Transformations between IFDS and IDE problems and their results}%
  \label{fig:transformations}%
\end{figure}

@@

\begin{example}
  Consider the program Listing~\ref{list:ccexample}. The exploded supergraph for that program is shown in Figure~\ref{fig:cc_edgefn_example}.

>>>>

An overview of the transformations is shown in Figure~\ref{fig:transformations}.

@@

\begin{example}
  Consider the program Listing~\ref{list:ccexample}. The exploded supergraph for that program is shown in Figure~\ref{fig:cc_edgefn_example}.\mtodo{Move Figure~\ref{fig:cc_edgefn_example} up to where it's mentioned first? If yes, the caption won't make sense.}
  \otodo{Can we combine the IDE Figure\ref{fig:cc_edgefn_example} with the IFDS Figure 4? It should be the same figure, just that Figure\ref{fig:cc_edgefn_example} has labels on the edges. We can initially tell the reader to ignore the labels until later.}

### 2014-12-05 content/intro.tex
restructuring

\subsection{Correlated Calls}

@@

\subsection{IFDS and IDE}
\otodo{This section should be moved to the Background section (and suitably merged with the text there).}
The IFDS framework is a precise and efficient algorithm for data-flow analysis. IFDS was developed in 1995 by T.\,Reps, S.\,Horwitz, and M.\,Sagiv at the University of Wisconsin and has been used to solve a variety of data-flow analysis problems~\cite{bodden2013spl,naeem2008typestate,DBLP:conf/birthday/KreikerRRSWY13,tripp2009taj}. The IFDS analysis is a version of the classic \textit{functional approach} to data-flow analysis proposed by M.\,Sharir and A.\,Pnueli~\cite{pnueli1981two}.

Given a data-flow problem that satisfies the restrictions of IFDS, the algorithm provides a \textit{context-sensitive} solution in polynomial time. In other data-flow algorithms not based on the functional approach, the result of the analysis at the entry of a procedure ``merges'' the incoming data obtained from all callers of the procedure. As a consequence, there is one global data-flow result computed at the end of the procedure. Context-sensitivity, however, allows an analysis to compute the data-flow result for a given procedure as a \textit{function} of the data-flow value at the start of the procedure. 
In other words, the analysis result for a procedure varies depending on where the procedure was called from. This significantly improves the precision of a data-flow analysis, which is why context-sensitivity is an important advantage of IFDS over classic data-flow algorithms.

Compared to IFDS, most data-flow analyses are either general but do not run in polynomial time~\cite{knoop1992interprocedural,pnueli1981two} or handle a very specific set of problems~\cite{knoop1993efficient}.

The IFDS algorithm is applicable to problems which can be expressed with data-flow functions that satisfy certain restrictions. \textit{Inter-procedural} flow functions specify how data flows from the invocation of a procedure to its start, and from the procedure's end back to its call site. \textit{Distributive} flow functions are those that distribute over the confluence operator. In the context of IFDS, the confluence operator is called meet, and it can be either union or intersection. The data-flow facts on which the analysis operates must be a \textit{finite} set $D$. Each flow function operates on a \textit{subset} of $D$ (for example, the set of variables in the program) which makes the domain of the flow functions the power set of $D$. We describe the IFDS restrictions in detail in Section~\ref{sec:suitableifds}.

The IDE framework is an expressive extension to IFDS that was created by the same authors in 1996.
The problems that IDE can solve include, but are not limited to, IFDS problems. Just as the IFDS algorithm, the IDE algorithm is suitable for data-flow analyses that can be encoded with inter-procedural, distributive flow functions. However, in IDE, the domain of the flow functions is not restricted to sets $D$ of data-flow facts. The IDE domain of a flow function consists of \textit{environments} that map data-flow facts from the set $D$ to lattice elements.

As an example, in a constant propagation problem, an IDE environment would map each variable to the (possibly) constant value that it is bound to. To illustrate the distinction between IFDS and IDE we could say that IFDS can find out which variables in a program are constants, whereas IDE can additionally retrieve the values of the constant variables.

\subsection{Paper Outline}

@@

The contributions of this work are:
\begin{itemize}
  \item A transformation from IFDS to IDE problems that considers correlated calls.
  \item An implementation in Scala of the correlated-calls transformation and the IDE algorithm which is based on the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.
\end{itemize}

  We prove that the solution to an IDE problem that considers correlated calls is more precise than the solution to the original IFDS problem.
  We also show that the correlated-calls analysis is sound, i.e. that it never considers concrete execution paths as infeasible.

@@

\mtodo{Reference appendix?}The remainder of this paper is organized as follows. In the next section, we describe the IFDS and IDE analyses in detail. In Section~\ref{chapter:cca} we present the correlated-calls analysis as a transformation of IFDS problems into a special kind of IDE problem. Section~\ref{sec:ccdatastr} describes an efficient representation of the data structures that are required to define a correlated-calls IDE transformation.
In Section~\ref{chapter:eval} we address some implementation aspects of the correlated-calls analysis and present an evaluation of its results. Section~\ref{chapter:concl} contains concluding remarks.

>>>>

\section{Motivating Example}

@@

@@

The contributions of this work are:
\begin{itemize}
  \item A transformation from IFDS to IDE problems that considers correlated calls.
  \item An implementation in Scala of the correlated-calls transformation and the IDE algorithm which is based on the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.
\end{itemize}\mtodo{Reference the figure here.}

\begin{figure}
  \centering
    \tikzset{
  ashadow/.style={opacity=.25, shadow xshift=0.07, shadow yshift=-0.07},
}
  \tikzstyle{problem}=[fill=greyblue,text width=2.3cm,rounded corners,font=\small,text=charcoal,drop shadow={ashadow, color=greyblue}]
  \tikzstyle{result}=[fill=bisque,text width=1.8cm,rounded corners,font=\small,text=charcoal,drop shadow={ashadow, color=greyblue}]
\begin{tikzpicture}
    \node [problem] (ifds) {IFDS problem};
    \node [problem] (equiv) [above right=0.9cm and .7\dist of ifds.east] {Equivalent IDE problem};
    \node [problem] (ccide)[below=3cm of equiv.west, anchor=west] {Correlated-calls IDE problem};
    \node [result] (equivres) [right=\dist of equiv.east] {Equivalence-IDE result};
    \node [result] (ccres) [below=3cm of equivres.west,anchor=west] {Correlated-calls result};
    \node [result] (ifdsres) [right=\dist of equivres.east] {IFDS result};
    \node [result] (improved) [right=\dist of ccres.east] {Improved IFDS result};
    \path[->] (equivres) edge node[above]{$\backEq$} (ifdsres);
    \path[->] (ccres) edge node[above] {$\backCC$} (improved);
    \path[->] (ccres) edge[out=30,in=240] node[above] {$\backEq$} (ifdsres);
    \path[->] (ifds) edge[out=40,in=190] node[above] {$\transEq$} (equiv);
    \path[->] (ifds) edge[out=-60,in=165] node[above] {$\transCC$} (ccide);
    \path[->] (equiv) edge node[above]{$\result$} (equivres);
    \path[->] (ccide) edge node[above]{$\result$} (ccres);
\end{tikzpicture}
  \caption{Transformations between IFDS and IDE problems and their results}%
  \label{fig:transformations}%
\end{figure}

  We prove that the solution to an IDE problem that considers correlated calls is more precise than the solution to the original IFDS problem.
  We also show that the correlated-calls analysis is sound, i.e. that it never considers concrete execution paths as infeasible.\mtodo{As discussed on the phone, we want to talk more about the results here.}

@@

The remainder of this paper is organized as follows. In the next section, we describe the IFDS and IDE analyses in detail. In Section~\ref{chapter:cca} we present the correlated-calls analysis as a transformation of IFDS problems into a special kind of IDE problem. Section~\ref{sec:ccdatastr} describes an efficient representation of the data structures that are required to define a correlated-calls IDE transformation.
In Section~\ref{chapter:eval} we address some implementation aspects of the correlated-calls analysis and present an evaluation of its results. Section~\ref{chapter:concl} contains concluding remarks. The proofs for the lemmas in the paper are provided in the Appendix.\mtodo{Reference appendix like this?}
